---
title: "LiftOver test"
output:
  pdf_document: default
  html_document: default
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
# Set up the environment
library(knitr)
opts_chunk$set(cache.path='cache/', fig.path='img/', cache=T, tidy=T, fig.keep='high', echo=T, dpi=100, warnings=F, message=F, comment=NA, warning=F, results='as.is', fig.width = 10, fig.height = 6)
options(stringsAsFactors = FALSE)
```

```{r}
library(GenomicRanges)
library(rtracklayer)
library(liftOver)
library(R.utils)
```

# liftover using R

```{r}
# Make a GRanges object from this sample range
gr <- GRanges(seqnames = Rle(c("chr3")),
                 ranges = IRanges(start = 195084601, end = 195195500))
# The chain file ("hg19ToHg38.over.chain") can be found at: 
# http://hgdownload.soe.ucsc.edu/goldenPath/hg19/liftOver/
path <- "http://hgdownload.soe.ucsc.edu/goldenPath/hg19/liftOver/hg19ToHg38.over.chain.gz"
download.file(path, "hg19ToHg38.over.chain.gz")
gunzip("hg19ToHg38.over.chain.gz", skip = TRUE)
chain = import.chain("hg19ToHg38.over.chain")
 
gr_lift = liftOver(gr, chain)
gr_lift[[1]]
```

This results in 21 ranges for the single given range. Note the start coordinate of the first range and the end coordinate of the last range: 195363872 and 195474797 


# liftover using the web client:

Navigate to the web based tool @ https://genome.ucsc.edu/cgi-bin/hgLiftOver and submit the following coordinates (same as in above R chunk) in the submission box: 

`chr3 195084601 195195500` (OR formatted as `chr3:195084601-195195500` also works as input.) 

Select "Human" in both the Original and New Genome dropdown menus, 
and select "Feb. 2009 (GRCh37/hg19)" for the Original Assembly dropdown 
and "Dec. 2013 (GRCh38/hg38)" for the New Assembly. Click Submit

The web tool returns a downloadable text file that should read: 

`chr3	195363872	195474797` (Or `chr3:195363872-195474797`, if it was submitted in the alternate format)

Note that these coordinates match first start coordinate and 
the last end coordinate from the 21 ranges produced by liftover in R. 


# liftOver in a terminal

Instructions are here: http://genometoolbox.blogspot.com/2013/11/install-liftover-locally-on-unix.html 
but also summarized below.

Run `liftOver` on a Mac by downloading it @ http://hgdownload.cse.ucsc.edu/admin/exe/macOSX.x86_64/ 
and scrolling down to and downloading the `liftOver` package. 
Once downloaded move it to your working directory, and modify the permissions 
to make executable (i.e. `chmod u+x liftOver`). Make an input file with the coordinates, (e.g make a file `gr.bed` containing one line, tab-separated: 

```
chr3    195084601   195195500
```

To run liftOver, the usage is:

```
liftOver gr.bed hg19ToHg38.over.chain gr_lift.bed gr_lift.unmapped
```

This returns the same results as the web tool. The file `gr_lift.bed` contains 

```
chr3    195363872       195474797
```

```{r session_info}
sessionInfo()
```


